### 2015-11-01 2015/Temporal Distinctiveness & Mixture Distributions/Manuscript Files/main.tex
changed folder name

\documentclass[a4paper, jou, natbib]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}

%\usepackage{lineno}

\title{Temporal Distinctiveness in Task Switching: Assessing the Mixture-Distribution Assumption}
\shorttitle{Mixture Distribution Assumption}
\author{James A. Grange}
\affiliation{School of Psychology, Keele University, UK}

\note{{\bf Word Count:} x,xxx (Main body, not including abstract or references.)}
% \note{{\bf Under Review---Please, no direct quotes.}}

% command to insert editing comments or track changes
\newcommand{\jg}[1]{\textcolor{blue}{$^{\textrm{}}${#1}}}

%command for R symbol
   \newcommand{\R}{R}

\authornote{Please address correspondence to James A. Grange, School of Psychology, Dorothy Hodgkin Building, Keele University, Keele, UK, ST5 5BG. Email: grange.jim@gmail.com. All raw data, analysis code, and model code are available to download at www.GitHub.com/JimGrange/MixtureDistribution}

\leftheader{Grange}

\abstract{.}

\keywords{Task switching; decay; interference; computational model}

\begin{document}
\maketitle

%\bibliography{References.bib}

\end{document}